---
title: "GLAM - Project"
author: "Olivier Grognuz"
date: "04/05/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(GGally)
library(corrplot)
```

We would like to model the effect of the pollutants and the meteorological conditions on the number of hospital admissions.
So we have 2 questions:
1. Effect of the pollutants `no2`, `so2`, `rsp`, `o3`
2. Effect of meteorological conds `temp` and `hum`
Response variable is `resp`. 

# EDA 

```{r}
polcsv <- read.csv("pollution.csv")

head(polcsv)

pol <- na.omit(polcsv)
dim(pol)

pol$month <- as.factor(pol$month)
pol$year <- as.factor(pol$year)
pol$weekday <- as.factor(pol$weekday)
str(pol)
```

```{r}
#Some subsets for easier use

pol.quants <- pol[,!names(pol) %in% c("day", "month", "year", "date", "weekday")]

pol.quants.log <- pol.quants
pol.quants.log[c("no2", "so2", "rsp", "o3")] <- log(pol.quants.log[c("no2", "so2", "rsp", "o3")])
names(pol.quants.log)[2] <- "log(no2)"
names(pol.quants.log)[3] <- "log(so2)"
names(pol.quants.log)[4] <- "log(rsp)"
names(pol.quants.log)[5] <- "log(o3)"


pol.meteo <- pol[,c("temp", "hum")]
pol.pol <- pol[,c("no2", "so2", "rsp", "o3")]
```

All
```{r}
ggpairs(pol[, !names(pol) %in% c("date")])
ggsave("figures/pairs/all.png", width=10, height=8)
```


Only quantitative variables:
```{r}
ggpairs(pol.quants)
ggsave("figures/pairs/quant.png", width=10, height=7)

ggpairs(pol.quants.log)
ggsave("figures/pairs/quant_log.png", width=10, height=7)
```

Quants + month
```{r}
ggpairs(pol[,!names(pol) %in% c("day", "year", "date", "weekday")], aes(color=month, alpha=0.4))
ggsave("figures/pairs/quant_month.png", width=9, height=6)
```

The `month` factor also explains in a way the meteorological conditions since summer is high temperature and winter low.

Quants + year
```{r}
ggpairs(pol[,!names(pol) %in% c("day", "month", "date", "weekday")], aes(color=year, alpha=0.4))
ggsave("figures/pairs/quant_year.png", width=9, height=6)
```

Quants + weekday
```{r}
ggpairs(pol[,!names(pol) %in% c("day", "month", "date", "year")], aes(color=weekday, alpha=0.4))
ggsave("figures/pairs/quant_weekday.png", width=9, height=6)
```

Correlation matrix
```{r}
cor(pol.quants)
corrplot(cor(pol.quants))
```

Only thing to notice is rsp <=> no2, does it mean something? 78% correlation

With the plots, we don't really see any clear pattern apart that it seems to increase with the years.
But, maybe we have an interaction effect! We should investigate that as well.
